---
title: "General Social Survey"
author: "Erika Munoz"
date: "October 26, 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
In my method I would be using two catigoricaL variables. For our graphical descriptive results we will use a bar chart. For our numerical results we will be using the xtab test. For our analysis we will be using the chisq test. 

$H_0$:"There is no relationship between the death penalty and gun ownership."

$H_A$:"There is a relationship between the death penalty and gun ownership."


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC (~CAPPUN+OWNGUN, data=mydata4, type="percent", xlab="capital punishment", main="Capital Punisghment VS. Gun Ownership")
```
The People that favor capital punishment there is about a 5% different between owning a gun and not owning a gun. The people that oppose capital punishment have a very big difference. There is more people that do not own guns rather than people that own guns. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab1=xtabs(~CAPPUN+OWNGUN, data=mydata4)
rowPerc(tab1)
```
The people that favor capital punishment have a 3.44% difference of the people that own guns and those that do not own guns. The people that oppose guns have a 41.94% different from those that own guns and those that do not own guns. 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisqtestGC(tab1)
```
We reject the null hypothesis $H_0$:"There is no relationship between the death penalty and gun ownership." therfore we are able to conclude that there is a relationship between capital punishment and gun ownership. According to our bar chart and xtabs results there is a big difference of gun ownership between the people that oppose capital punishment. Rather than the people that are for capital punishment that have a slight difference of gun ownership. 


# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG)  independent of belief in the afterlife (POSTLIFE)


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

In my method I would be using two catigoricaL variables. For our graphical descriptive results we will use a bar chart. For our numerical results we will be using the xtab test. For our analysis we will be using the chisq test. 

$H_0$:"There is no relationship between the belief of the  Bigbang theory and the afterlife theory."

$H_A$:"There is a relationship between the belief of the Bigbang theory and the afterlife theory."


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, type="percent", xlab="BIGBANG", main="Bigbang VS. The After Life")
```

There is no difference between the people that believe and do not believe in the bigbang theory. According to the graph most of the people belive in the after life. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab2=xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(tab2)
```
The people that believe in if the bigbang theory is true have a 81.71% of believe in the after life and the people that do not believe in the bigbang theory have a 91.67% of believe in the afterlife. 

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab2)

```
We fail to reject the null hypothesis $H_0$:"There is no relationship between the belief of the  Bigbang theory and the afterlife theory." therfore that statment is true. Our alternitive hypothesis $H_A$:"There is a relationship between the belief of the Bigbang theory and the afterlife theory." is therfore false.

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

In my method I would be using two catigoricaL variables. For our graphical descriptive results we will use a bar chart. For our numerical results we will be using the xtab test. For our analysis we will be using the chisq test.

$H_0$:"There is no relationship between a persons opinion on the death penalty and their race."

$H_A$:"There is a relationship between a persons opinion on the death penalty and their race."

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+RACE, data=mydata4, type="percent", xlab="Capital Punishment", main="Capital Punishment VS. Race")
```
In this graph the people that favor capital punishment are a majority of white and other. The people that oppose capital punishment are a majority of black people. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
tab3=xtabs(~CAPPUN+RACE, data=mydata4)
rowPerc(tab3)
```
There is a 79.31% of White people that favor capital punishment compared to the 20.69% that were a different race. However there is a favor in opposition amongst black race as they have a 27% to oppose capital punishment.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(tab3)

```
We reject the null hypothesis $H_0$:"There is no relationship between a persons opinion on the death penalty and their race." therfore there is a relationship between the death penalty and race. There is a majority of White race that are for capital punishment while the black race oppose capital punishment. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In questions one and three the variables do effect there relationship. But on question two there is no relationship between the two variables. 